```
---
title: "CCD"
author: "William C Nelson"
date: "3/28/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
Load necessary packages
```{r}
library(pheatmap)
library(ISLR)
library(ggplot2)
library(GGally)
library(ggfortify)
library(umap)
library(dplyr)
library(reshape2)
library(stringr) # for str_detect()
```
Load in the image from the morgannprice.org project page
```{r}

load("/Users/shara/Desktop/Mission_Sakura/Phase_I_SBW25_RbTnSeq_FirstStrainRecommendations/RbTnSeq_data/other_data/fit.image")

```
Get the unique experimental conditions
```{r}
# Get the unique conditions 
conditions<-unique(c(unique(expsUsed$Condition_1), unique(expsUsed$Condition_2), unique(expsUsed$Condition_3) ))
conditions<-conditions[!conditions %in% c("", NA)]
conditions<-conditions[!str_detect(conditions,"Soil")]
# Get the unique pH
pHvals<-unique(expsUsed$pH)
pHvals<-pHvals[!pHvals %in% c("", NA)]
# Get the unique Temperatures
Tinput<-unique(expsUsed$Temperature)
Tinput<-Tinput[!Tinput %in% c("", NA)]
# Creating a nested list of all the conditions
```
all_criteria_names <- c("compound","pH","Temperature")
all_criteria <-list()
all_criteria[["compound"]] <- conditions
all_criteria[["pH"]] <- pHvals
all_criteria[["Temperature"]] <- Tinput
```
all_criteria<- list("Condition" = conditions, "pH" = pHvals, "Temperature" = Tinput)

```
Iterate through all the conditions - This is not working!!!
```{r}
for (parameter in names(all_criteria))
{
  for (i in 1:length(all_criteria[[parameter]]))
  {
    print(paste('i=',i))
    # --------------------------------------------------------------------------
    # Value of the experimental condition is obtained
    param_val = all_criteria[[parameter]][i]
    # --------------------------------------------------------------------------
    # All the corresponding experiment samples corresponding to the parameter is obtained
    if (parameter == "Condition")
    {
      # Search across Condition_1, Condition_2 and Condition_3
      Exps<-expsUsed[which((expsUsed$Condition_1 ==param_val) | (expsUsed$Condition_2==param_val) | (expsUsed$Condition_3==param_val)),]
    }else
    {
      Exps<-expsUsed[which(expsUsed[parameter] == param_val),]
    }
    # --------------------------------------------------------------------------
    # The number of experiments should be greater than 20 to proceed
    if (nrow(Exps)<3){next} #-------------------------------------------------------------has been commented-------------------------------------------------------------------------------------------------------------------------
    # --------------------------------------------------------------------------
    # Conditional Cofitness Analysis
    condCofit<-cor(t(fit$lrn[,fit$q$u & names(fit$lrn) %in% Exps$name]))
    # Subtract the identity matrix to remove the diagonal self correlation entries which are basically 1
    condCofit<- condCofit - diag(dim(condCofit)[1])
    # Need to figue out what "genCofit" means -----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
    genCofit<-matrix(0,dim(condCofit)[1],dim(condCofit)[2])
    CCD<-condCofit - genCofit
    condCofit_cutoff = 1.0
    repeat{
      condCofit_cutoff<-condCofit_cutoff - 0.1
      if(condCofit_cutoff<0.5){
      break
      }
      screen<- which((abs(condCofit) >= condCofit_cutoff), arr.ind=T)
      if(nrow(screen)<12000 & nrow(screen)>8000){
        break
      }
    }
    fitGenes_cofit<-unique(c(fit$g[screen[, "col"]], fit$g[screen[, "row"]]))
    # --------------------------------------------------------------------------
    # Find the average log ratio normalized(lrn) and average t-statistic
    if(sum(fit$q$u & colnames(fit$lrn) %in% Exps$name)<3){
    next
    }
    avgLRN<-apply(fit$lrn[,(fit$q$u & colnames(fit$lrn) %in% Exps$name)], 1, mean)
    avgT<-apply(fit$t[,(fit$q$u & colnames(fit$lrn) %in% Exps$name)], 1, mean)
    # --------------------------------------------------------------------------
    # Filter the genes based on criteria
    # Condition 1 : Genes should have a condCofit > condCofit_cutoff
    # Condition 2 : avgLRN(gene) > 2
    # Condition 3 : avgT(gene) > 5
    fitGenes<-fit$g[abs(avgLRN) >= 2 & abs(avgT) >= 5]
    fitGenes<-fitGenes[fitGenes %in% fitGenes_cofit]
    # --------------------------------------------------------------------------
    # Print the summary of what happened in the current run
    print('++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++')
    if (parameter == "Condition")
    {
      print(paste('Condition : ', param_val))
      condition_name = paste("compound:",param_val,sep="")
    }
    else
    {
      print(paste(parameter, ':', param_val))
      condition_name = paste(parameter,":",param_val,sep="")
    }
    print(paste('Number of experiments :', dim(Exps)[0]))
    print(paste('Number of identified genes :', length(fitGenes)))
    write.table(fitGenes,paste('/Users/Shara/Box/YeungLabUCSBShare/Shara/_Project_Naruto/Phase_I_SBW25_RbTnSeq_FirstStrainRecommendations/gene_predictions/',condition_name,".csv",sep=""),sep='\t', col.names=T)
  }
}
print('++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++')


# I'm supposedly done here

# condition_name --- to be defined prior to exporting the data
# write.table(fitGenes,paste('/Users/Shara/Box/YeungLabUCSBShare/Shara/_Project_Naruto/Phase_I_SBW25_RbTnSeq_FirstStrainRecommendations/gene_predictions/',condition_name,".csv",sep=""),sep='\t', col.names=T)





#===============================================================================
#===============================================================================
#===============================================================================

i=10
param_val = all_criteria[[parameter]][i]
Exps<-expsUsed[which(expsUsed$Condition_1 ==param_val) || which(expsUsed$Condition_2==param_val) || which(expsUsed$Condition_3==param_val),]
dim(Exps)
#===============================================================================
#===============================================================================
```
Load the experiments
```{r}
allExpsUsed=read.csv(file="expsUsed.csv")
expsUsed=expsUsed[expsUsed$Date_pool_expt_started=="September 1, 2021",]
```

Define necessary variables
```{r}
condition="Sodium Chloride"
#pH=8.5
#temp=8
#condition = "L-Glutamine"
#condition = "Betaine"
#condition = "Methyl 3-(4-hydroxyphenyl)propionate" # not enough experiments.
#condition = "D,L-Malic Acid"
#condition = "p-Coumaric acid"
```

Pull the experiments for the specific condition
```{r}
Exps<-expsUsed[expsUsed$Condition_3 == condition,]
#Exps<-expsUsed[expsUsed$Condition_1 == condition,]
#Exps<-expsUsed[expsUsed$Temperature == temp,]
#Exps<-expsUsed[expsUsed$pH == pH,]
nrow(Exps)
```
That number above should be higher than 20 to continue.

Filter for CCD results for genes with good fitness values under the specific
condition.
```{r}
# calculate the avg LRN  and t for each gene across the target condition
avgLRN<-apply(fit$lrn[,(fit$q$u & colnames(fit$lrn) %in% Exps$name)], 1, mean)
avgT<-apply(fit$t[,(fit$q$u & colnames(fit$lrn) %in% Exps$name)], 1, mean)
# select only those genes that beat the cutoffs
fitGenes<-fit$g[abs(avgLRN) >= 2 & abs(avgT) >= 5]
length(fitGenes)
```
The number above should be the number of genes showing significant fitness effects
under the give growth condition (|LRN|>2, |t|>5)

Calculate the conditional cofitness
```{r}
condCofit<-cor(t(fit$lrn[,fit$q$u & names(fit$lrn) %in% Exps$name]))
hist(condCofit, breaks=200)
quantile(condCofit, probs=c(0.001, 0.25, 0.5, 0.75, 0.999))
```
If this curve is flat, that means there is not enough data to distinguish between conditions, and going further may not be fruitful.

Calculate the cofitness difference
```{r}
CCD<-condCofit - genCofit
hist(CCD, breaks=200)
hist(CCD/max(abs(CCD)), breaks=200) # this histogram is more comparable to the one above.
```

Screen for the best/beat cutoff (the same gene pairs screened for their 
specific condition cofitness)
```{r}
# find the best cofitness value for each gene (that is not 1, which should be the self-cofitness)
#bestcofit<-apply(condCofit, 1, function(x) min(-abs(x[x!=1])))
# find which pairs have either an absolute cofitness >0.7 or are the best cofitness value
#screen<- which(-abs(condCofit) == bestcofit | (-abs(condCofit) <= -0.7 & condCofit != 1), arr.ind=T)
screen<- which((-abs(condCofit) <= -0.7 & condCofit != 1), arr.ind=T)
#screen<- which((-abs(CCD/max(abs(CCD))) <= -0.7), arr.ind=T)
#screen<- which((-abs(condCofit) <= -0.9 & condCofit != 1), arr.ind=T)
nrow(screen)
```
What is a good number for the number of screen rows? 10k seems reasonable

```{r}
out = matrix(nrow=0, ncol=11)
for (i in 1:nrow(screen)) {
  # make sure we're only pulling one comparison
  if (screen[i,"row"] > screen[i, "col"]) {
    targetGene<-fit$g[screen[i, "col"]]
    hitGene<-fit$g[screen[i, "row"]]
    CCFVal<-condCofit[screen[i,"row"], screen[i, "col"]]
    CCDVal<-CCD[screen[i,"row"], screen[i, "col"]]
    GCFVal<-genCofit[screen[i,"row"], screen[i, "col"]]
    targetLRNVal<-avgLRN[screen[i, "col"]]
    hitLRNVal<-avgLRN[screen[i,"row"]]
    targetTVal<-avgT[screen[i, "col"]]
    hitTVal<-avgT[screen[i,"row"]]
    out<-rbind(out, c(targetGene, hitGene, CCFVal, GCFVal, CCDVal, targetLRNVal, targetTVal, hitLRNVal, hitTVal, SBW25[targetGene, 'description'], SBW25[hitGene,'description']))
  }
}
# because I haven't learned how to autotype dataframes...
out<-as.data.frame(out)
out$V3<-as.numeric(out$V3)
out$V4<-as.numeric(out$V4)
out$V5<-as.numeric(out$V5)
colnames(out)<-c("gene1", "gene2","condCofit", "genCofit", "CCD", "gene1LRN", "gene1t", "gene2LRN", "gene2t", "gene1desc", "gene2desc")
```


```{r}
# grab any rows from CCD out which have one of the fitGenes as either gene
fitGenesCCDout<-out[out$gene1 %in% fitGenes | out$gene2 %in% fitGenes,]

# this is our output
write.csv(fitGenesCCDout, file=paste(condition,"_CCD.csv"), quote=F)

pdf(file=paste(condition, "_CCD.pdf"))
# look at the heatmap for those pairs where the CCD beats cutoff
pheatmap(fitGenesCCDout[,3:5], cluster_cols=F, cellwidth=25, show_rownames=F)
#pheatmap(fitGenesCCDout[abs(fitGenesCCDout[,5])>0.7,3:5], cluster_cols=F, cell_height=3, cell_width=20, show_rownames=F)
dev.off()
```
We want output that is every gene that confers fitness above cutoffs (fitGenes) and any CCD that beats cutoff (what is cutoff?). Return a number of relationships = 1/1000 the total matrix (so ~30k)
Look at the 0.001 and .999 quantiles of condCofit and averag their abs vals. If number is <0.7, use 0.7.

```{r}
out = matrix(nrow=0, ncol=11)
fitGeneRows<-which(row.names(CCD) %in% fitGenes)
thisCutoff=0.7
for (rowID in fitGeneRows) {
  screenRows<-which(abs(condCofit[rowID,]) > thisCutoff)
  for (colID in screenRows) {
    targetGene<-fit$g[rowID]
    hitGene<-fit$g[colID]
    CCFVal<-condCofit[rowID,colID]
    CCDVal<-CCD[rowID, colID]
    GCFVal<-genCofit[rowID, colID]
    targetLRNVal<-avgLRN[rowID]
    hitLRNVal<-avgLRN[colID]
    targetTVal<-avgT[rowID]
    hitTVal<-avgT[colID]
    out<-rbind(out, c(targetGene, hitGene, CCFVal, GCFVal, CCDVal, targetLRNVal, targetTVal, hitLRNVal, hitTVal, SBW25[targetGene, 'description'], SBW25[hitGene,'description']))

  }
}
out<-as.data.frame(out)
out[,3:9]<-sapply(out[,3:9],as.numeric)
colnames(out)<-c("gene1", "gene2","condCofit", "genCofit", "CCD", "gene1LRN", "gene1t", "gene2LRN", "gene2t", "gene1desc", "gene2desc")

# Choose the right output file name formatting
write.csv(out, file=paste(condition,"_CCD2.csv"), quote=F)
#write.csv(out[(abs(out$CCD)>=thisCutoff/3) | (out$genCofit==1),], file=paste(temp,"C_CCD2.csv"), quote=F)
#write.csv(out[(abs(out$CCD)>=thisCutoff/3) | (out$genCofit==1),], file=paste("pH", pH,"_CCD2.csv"), quote=F)

#pdf(file=paste(condition, "_CCD2.pdf"))
# look at the heatmap for those pairs where the CCD beats cutoff
pheatmap(out[,3:5], cluster_cols=F, cellwidth=25, show_rownames=F)
#pheatmap(fitGenesCCDout[abs(fitGenesCCDout[,5])>0.7,3:5], cluster_cols=F, cell_height=3, cell_width=20, show_rownames=F)
#dev.off()

```

